---
title: "Exploring College Score Card With R"
author: "Parsa Shams"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'), 
                      echo=F, 
                      cache=T, 
                      warning = F,
                      message=F)
options(htmltools.dir.version = FALSE)
library(rscorecard)
library(dplyr)
library(magrittr)
library(ggmap)
library(ggplot2)
library(zipcode)
library(ggthemes)
source("config.R")
```

# Motivation

`r newthought("This document is the confluence")` of two curiousities. First, I have been wanting to explore the College Score Card dataset, an open dataset curated by US Department of Education, which provides a wealth of information about American degree granting colleges and universities. Second reason is the desire to test the intruiging elegance of Edward Tufte's Handout Style of visual information design. Currently, I have no specific plans for this project. I will only be able to devote scattered increments of time to this exploration. The goal is to just explore a new dataset in a new medium and see if the medium is condusive to clearer data-driven story telling. 

```{r get-data, include=FALSE}
df <- sc_init() %>% 
    sc_filter(PREDDEG==3) %>%
    sc_select(unitid, instnm, PREDDEG, zip, 
              ADM_RATE, PCTPELL, C150_4, C150_L4, CCBASIC,CONTROL) %>% 
    sc_get()

data(zipcode)

df %<>%
  mutate(zip = clean.zipcodes(zip)) %>%
  inner_join(zipcode,by="zip")

base_map <- get_map(location='united states', zoom=4 , maptype = "toner-lite",color='color')
```

# Graduation Rate


```{r, fig.width = 10, fig.height=3.5,fig.fullwidth=TRUE, cache=TRUE,fig.cap="4-Year Graduation In US States"}
plt_df <- df %>% group_by(state) %>% 
  mutate(sort_x = -mean(c150_4,na.rm=T),
         n = n()) %>%
  filter(n>=10)

ggplot(plt_df, aes(x=reorder(state,sort_x),y=c150_4*100)) +
  stat_summary(fun.data="mean_cl_boot") +
  # coord_flip() +
  theme_tufte() +
  geom_hline(yintercept=50) +
  labs(x="State",y="Graduation Rate") 
  # coord_flip()
```

The average graduation rate in the US is `r mean(df$c150_4,na.rm=T)`

let's map graduation rates on a map of states

# Selectiveness

```{r make-admit-map, cache=TRUE, message=FALSE, warning=FALSE}
ggmap(base_map, base_layer=ggplot(df, aes(x=longitude, y=latitude))) + 
geom_point(
      aes(colour=adm_rate),
      alpha=.5)  +
scale_color_distiller(palette="Spectral") +
theme_void()
```




# Paying for College

```{r, fig.height=5, fig.width=8, fig.fullwidth=TRUE}
left <- 
  ggplot(df,aes(x=is.na(adm_rate),y=c150_4)) +
  geom_boxplot() +
  scale_x_discrete("Admit Rates",labels=c("Missing","Reported")) +
  labs(y="Graduation Rate") +
  theme_tufte()


right <- 
  ggplot(df) +
  geom_point(aes(adm_rate,c150_4,color=pctpell)) +
  scale_color_distiller(palette="Spectral") +
  labs(x="Admit Rate", y="4-year Graduation Rate",color="Pell Eligible")+
  coord_fixed(ratio=1) +
  theme_tufte()


gridExtra::grid.arrange(left,right,ncol=2)
```